Concentration Inequalities

### 2015-09-29 causal-bandit/observe_then_best.tex
adds lots of figures for TPR
updated causal bandits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{times}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{comment}
\usepackage{amsthm}
\usepackage{natbib}
\usepackage[capitalize]{cleveref}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{tikz} 
\usetikzlibrary{arrows,positioning} 
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}
\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}
\newcommand{\eqn}[1]{\begin{align}#1\end{align}}
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\Ber}{\operatorname{Bernoulli}}
\renewcommand{\P}[1]{\operatorname{P}\left\{#1\right\}}

@@

@@

For a fixed $m$ the $\boldsymbol{q}$ most likely to lead us to overestimate $m$ is:

@@

For $i < m$,we have $\bar{q}_i = q_i = 0$. For $i \geq m$
\eq{
P(q_i - \bar{q}_i \geq C) & =  P(q_i - \hat{q}_i \geq C|\hat{q}_i\leq \frac{1}{2})P(\hat{q}_i \leq \frac{1}{2})+P(q_i - (1-\hat{q}_i) \geq C | \hat{q}_i > \frac{1}{2})P(\hat{q}_i>\frac{1}{2})\\
& \leq  2P(q_i - \hat{q}_i \geq C|\hat{q}_i\leq \frac{1}{2})
}

Via Bernstein's Inequality:

\eqn{
&2P(q_i - \hat{q}_i \geq C | \hat{q}_i \leq \frac{1}{2}) \leq 2\exp(-\frac{hC^2}{4q_i}) := \gamma \\
\implies &P(q_i - \hat{q}_i \geq 2\sqrt{\frac{\log(2/\gamma)}{mh}})\leq \gamma
}

>>>>

\usepackage{todonotes}

%\usepackage[disable]{todonotes}
\newcommand{\todot}[2][]{\todo[color=red!20!white,#1]{#2}}
\newcommand{\todof}[2][]{\todo[color=orange!20!white,#1]{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{times}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{comment}
\usepackage{amsthm}
\usepackage{natbib}
\usepackage[capitalize]{cleveref}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{tikz} 
\usetikzlibrary{arrows,positioning} 
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}
\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}
\newcommand{\eqn}[1]{\begin{align}#1\end{align}}
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\Ber}{\operatorname{Bernoulli}}
\renewcommand{\P}[1]{\operatorname{P}\left\{#1\right\}}

@@

\section{Bounding the Simple Regret}

\begin{itemize}
\item $N$ is number of variables.
\item $q_i$ is probability that $X_{i,t} = 1$ for any $t$
\item $s_i = \min\set{q_i, 1 - q_i}$
\item $h = T/2$
\item $\mu : \set{1,\ldots,N} \times \set{0,1} \to [0,1]$.
\item $I_t \in \set{1,\ldots,N} \times \set{0,1}$ is the intervention
\end{itemize}

Let $r_T$ be the simple regret after $T$ time steps, which is
\eq{
r_q(T) = \E\left[\mu^* - \mu_{I_T}\right]\,,
}
where $\mu^*$ is the mean payoff of the optimal intervention and $I_T$ is the intervention chosen in the $T$th round.
We propose a two part algorithm that is provably order optimal when its performance is measured in terms of the simple regret.

\begin{algorithm}[H]
\caption{Simple Regret Algorithm}\label{alg:simple}
\begin{algorithmic}[1]
\State {\bf Input:} $T, N$
\State $h = T/2$
\For{$t \in 1,\ldots,h$}
\State Do nothing and observe $X_{i,t}$ for $i \in \set{1,\ldots,N}$ and $r_t$
\EndFor
\State Compute $\hat q_i = \frac{1}{h} \sum_{t=1}^n X_{i,t}$
\State Compute $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$
\State Compute $\hat m = \min\set{2 \leq i \leq N : \hat s_{(i)} \geq 1/i}$
\State Compute $A$ is subset of insufficiently explored arms with $|A| = \hat m$
\State Compute for all $i \in \set{1,\ldots,N}$ and $j \in \set{0,1}$:
\eq{
\hat \mu_{i,j} = \frac{\sum_{t=1}^h \ind{X_i = j} r_t}{\sum_{t=1}^h \ind{X_i = j}}\,.
}
\For{$i \in A$}
\State $j = \ind{\hat q_{I_t} \geq 1/2}$
\For{$t \in 1,\ldots,h/\hat m$}
\State Choose $I = i$ and $J = j$
\State Observe reward $r_t$
\EndFor
\State Recompute $\hat \mu_{i,j} = \frac{\hat m}{h} \sum_{t=1}^{\hat m/h} r_t$ 
\EndFor
\end{algorithmic}
\end{algorithm}

\begin{theorem}\label{thm:simple-regret}
Define $m = \min\set{i : q_i \geq 1/i}$.
Then the algorithm given in \cref{alg:simple} satisfies
\eq{
r_q(T) \in O\left(\sqrt{\frac{m}{T} \log \left(\frac{NT}{m}\right)}\right)\,.
}
\end{theorem}

\begin{lemma}\label{lem:conc1}
$\displaystyle \P{\left|\hat q_i - q_i\right| \geq \sqrt{\frac{3q_i}{h} \log \frac{2}{\delta}}} \leq \delta$.
\end{lemma}

\begin{proof}
Let $Z_t = \ind{X_{i,t} = 1} \in \set{0,1}$.
Then
\eq{
\hat q_i = \frac{1}{h} \sum_{t=1}^{h} Z_t\,.
}
Now $Z_1,\ldots,Z_h$ is an i.i.d.\ sequence of Bernoulli random variables with with mean $q_i$. The result follows from the Chernoff bound.
\end{proof}

\begin{lemma}\label{lem:conc2}
Let $\delta >0$ and $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$ and
define $\hat m = \min\set{2 \leq i \leq N : \hat s_{(i)} \geq 1/i}$.
If
\eq{
h \geq 12m \log\frac{2N}{\delta}\,,
}
then
\eq{
\P{\hat m \leq 2m} \geq 1 - \delta\,.
}
\end{lemma}

\begin{proof}
It is easy to see that the worst case occurs when 
\eq{
q_i = \begin{cases}
0 & \text{if } i < m \\
\frac{1}{m} & \text{otherwise}\,.
\end{cases}
}
Now by \cref{lem:conc1} we have with probability at least $1 - \delta$ that 
\eq{
(\forall i) \qquad \left| \hat q_i - q_i\right| 
&\leq \sqrt{\frac{3q_i}{h} \log \frac{2N}{\delta}} \\
&= \sqrt{\frac{3}{mh} \log\frac{2N}{\delta}}\,. 
}
By assumption 
\eq{
h \geq 12m \log \frac{2N}{\delta}\,.
}
Therefore with probability at least $1 - \delta$ we have
\eq{
(\forall i) \qquad \left|\hat q_i - q_i\right| \leq \frac{1}{2m}\,.
}
Note that for $i$ with $q_i = 0$ it is easy to see that $\hat q_i = 0$ (the variance is zero). The above guarantee that
for $i$ with $q_i = 1/m$ we have $\hat q_i \in [1/(2m), 3/(2m)]$ for all $i$ with probability at least $1 - \delta$,
which implies that $\hat s_i \geq 1/(2m)$ for all $i$ where $q_i = 1/m$.
Therefore $\hat m \leq 2m$ with probability at least $1 - \delta$. 
\end{proof}

\begin{proof}[Proof of \cref{thm:simple-regret}]
Given the assumption that
\eq{
h \geq 12m \log \frac{N}{\delta}
}
we have shown with probability at least $1 - \delta$ that
\eq{
(\forall i, j) \qquad \left|\hat \mu_{i,j} - \mu_{i,j}\right| \leq \sqrt{\frac{m}{h} \log\frac{N}{\delta}}\,.
}
Suppose $h < 12m \log \frac{N}{\delta}$. Then
\eq{
\left|\hat \mu_{i,j} - \mu_{i,j}\right| \leq 1 \leq \sqrt{\frac{12m}{h} \log \frac{N}{\delta}}\,.
}
Therefore
\eq{
\left|\hat \mu_{i,j} - \mu_{i,j}\right| \leq \sqrt{\frac{12m}{h} \log \frac{N}{\delta}}\,.
}
Therefore 
\eq{
r_q(T) 
&\leq \delta + \sqrt{\frac{12m}{h} \log \frac{N}{\delta}} \\
&= \delta + \sqrt{\frac{24m}{T} \log \frac{N}{\delta}} \\
&\leq \frac{m}{T} + \sqrt{\frac{24m}{T} \log \left(\frac{NT}{m}\right)}
}
as required.
\end{proof}

@@

For a fixed $m$ the $\boldsymbol{q}$ most likely to lead us to overestimate $m$ is: \todot{Justify this claim.}

@@

For $i < m$,we have $\bar{q}_i = q_i = 0$. For $i \geq m$ \todot{Is this really true, justify}
\eq{
P(q_i - \bar{q}_i \geq C) & =  P(q_i - \hat{q}_i \geq C|\hat{q}_i\leq \frac{1}{2})P(\hat{q}_i \leq \frac{1}{2})+P(q_i - (1-\hat{q}_i) \geq C | \hat{q}_i > \frac{1}{2})P(\hat{q}_i>\frac{1}{2})\\
& \leq  2P(q_i - \hat{q}_i \geq C|\hat{q}_i\leq \frac{1}{2})
}

Via Bernstein's Inequality:
\eqn{
&2P(q_i - \hat{q}_i \geq C | \hat{q}_i \leq \frac{1}{2}) \leq 2\exp(-\frac{hC^2}{4q_i}) := \gamma \\
\implies &P(q_i - \hat{q}_i \geq 2\sqrt{\frac{\log(2/\gamma)}{mh}})\leq \gamma
}
